Select-based random access to variable-byte encodings

### 2020-02-28 gradu/msc_english_contents.tex
Changes

A variable-byte encoding based integer compression method with fast random access was first introduced by \citep{Bri09}. They used clever block 
reorganizing and $rank$ data structure to achieve random access. Their solution is currently the only published solution for the problem and it 
has been widely adopted.

An alternative solution with $select$ data structure is shown later. (TODO: refer to results)

--contributions?

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{bris.png}
\caption{Data structure by Brisaboa et al., visualized} \label{bris_ds}
\end{figure}

For example, if an element needed 3 chunks when encoded, the least significant bits would go to $A_1$, second least significant bits to $A_2$ and the most significant 
bits would go to $A_3$. When fetching i-th element, getting the first chunk is just fetching $A_1$[i]. If the continuation bit is set at 0 (meaning there are more chunks 
to this element), getting the correct index for $A_2$ is calculated with $rank_0$($A_1$, i). A pseudo code example of DAC with $rank$ is shown in Figure~\ref{bris_pseudo} 
with block length of 8. (TODO: example needed?)

@@

\draw[arrow] (Btarget.south) to (Atarget.north);
\end{tikzpicture}
\caption{Data} \label{VBarrays}
\end{figure}
TODO: try bitvector SD performance + size (select)

>>>>

A variable-byte encoding based integer compression method with constant time random access was first introduced by \citep{Bri09}. They used clever block 
reorganizing and $rank$ data structure to achieve random access. Their solution is currently the only published solution for the problem and it 
has been widely adopted.

An alternative solution with $select$ data structure is explained in detail and compared to $rank$ by Brisaboa et al. This method does not use data 
reorganizing, but rather capitalizes on the assumption that data is encoded the same way it's in the source. Because variable-byte encoding does not assign 
new codewords to data when encoding, reading the data straight from the memory is much faster than reassembling the word from blocks.

(TODO: refer to results)

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{bris.png}
\caption{Data structure by Brisaboa et al., visualized} (TODO: create this with tikz)\label{bris_ds}
\end{figure}

For example, if an element needed 3 chunks when encoded, the least significant bits would go to $A_1$, second least significant bits to $A_2$ and the most significant 
bits would go to $A_3$. When fetching i-th element, getting the first chunk is just fetching $A_1$[i]. If the continuation bit is set at 0 (meaning there are more chunks 
to this element), getting the correct index for $A_2$ is calculated with $rank_0$($A_1$, i). A pseudo code example of DAC with $rank$ is shown in Figure~\ref{bris_pseudo} 
with block length of 8.

@@

\draw[arrow] (Btarget.south) to (Atarget.north);
\end{tikzpicture}
\caption{Data} \label{VBarrays}
\end{figure}
TODO: try bitvector SD performance + size (select) (maybe write about different select implementations)
